%-----------paper summary---------------------
% answer the question asked by this study 
We have evaluated the potential for using observations of Earth rotation parameters to evaluate data assimilation systems and constrain atmospheric models as assimilated variables.  
% a high-level summary of what my data show:
% (1) ER data is a convenient way to evaluate how close an analysis is getting to the truth
% (2) But if we want to assimilate ER data, we find that the covariance model doesn't work right because it is trying to capture tiny correlations 
% (2) if we constrain the ensemble more using other obs, it still doesn't really work because we end up inhibiting out ability to fit the conventional obs by insisting that it fit these small, flawed correlations. 
%
Because they are freely available at high accuracy and relatively high temporal resolution and updated continuously \citep{iers}, 
Earth rotation parameters are useful as simple observables against which to validate the constraint imposed by observations in a data assimilation system.  
Comparing the angular momentum of a model ensemble to these observations is computationally straightforward;
for a standalone atmosphere model, component $\chi_2$ is the most useful because it is the most dominated by the atmosphere and can be easily related to polar motion without an ambigious constant of integration. 
This comparison gives a global view both of how well the modeled wind and mass fields approximate the truth.
The Earth rotation parameters have been connected to various atmospheric models, and their oceanic residuals have even been used to constrain an ocean model via data assimilation \citep{Saynisch2010,Saynisch2011,Saynisch2012}. 

However, assimilation experiments with synthetic observations and a known truth showed that it is difficult to constrain the model state with angular momentum observations because they are global integrals of the model fields, which leads to 
associated state-to-observation covariances that are much smaller than any realizible ensemble sampling error (Fig. \ref{fig:covariances}). 
Thus the ensemble filter used in this study generates a covariance model that is physically plausible and computes the statistically most likely update of each state space component, but because of sampling error it is common that 
the state update computed by the filter moves the ensemble and its mean farther away from the true state, while still satisfying the angular momentum observations (red contours in Fig. \ref{fig:error_increments}). 
Though the global angular momentum observations can impose a weak constraint on some aspects of the modeled state (Fig. \ref{fig:point_checks}), it is not enough to yield a reliable and robust correction to the model state. 

One might therefore expect that global angular momentum observations can constrain the modeled state better if the ensemble is already largely constrained to the truth by more conventional, spatially-localized observations. 
We tested this idea by assimilating a global grid of synthetic temperature observations in addition to the global angular momentum observations. 
We found that, overall, adding the angular momentum observations worsens the analysis (Fig. \ref{fig:added_value_MSE}c), 
because the sampling error inherent in the angular momentum assimilation inibits error reduction in the analysis increment. 

% anticipation of any criticisms and caveats  
A common technique for avoiding divergence of an ensemble filter is to periodically inflate the ensemble about its mean. 
It is unlikely that ensemble inflation would help here, since we found that the filter diverges both when the ensemble spread is large (section \ref{sec:erpda}) and when it is very small (section \ref{sec:added_value}). 
We have seen in a set of short assimilation runs with ensemble inflation (omitted here for brevity), that this is indeed the case. 


% my result in the context of what others have done  
Even though this study focused on an atmosphere model, the results can likely be extended to ocean models, where the state vector, although smaller, is still many orders of magnitude larger than the observation vector, and this would also yield state-to-observation correlations that are much smaller than the sampling error. 
For the ocean, it is possible to adjust the freshwater flux into the model, rather than the prognostic state variables, as a control parameter (as in \citet{Saynisch2010} and \citet{Saynisch2012}), but this has no corollary in atmosphere models. 


% points for future research  
Overall, this work illustrates the difficulties of assimilating observations that represent integrals or averages of the model state. 
\citet{Dirren2005} proposed a variation of the ensemble filter to deal with observations that are temporal averages of the state; their alternative algorithm projects the observation increment only on the time-average of each ensemble member while keeping deviations from the average untouched. 
\citet{Huntley2009} showed that applying this method in a global atmosphere model reduces errors over a wide range of scales, and that one can even improve instantanous errors when only time-average observations are assimilated. 
It is conceivable that this approach could be extended to the assimilation of Earth rotation parameters, by updating the global-average contribution to the angular momentum at each gridpoint, and thereby improve the state estimate. 
However, for assimilation systems on the scale of DART-CAM and DART-WACCM, this would require major changes in the code structure, while the possible gains of such a step are not clear. 


Another alternative pathway may be to assimilate the rate-of-change of Earth rotation parameters, thereby constraining the total torque between the atmosphere and the solid Earth. 
The net torque is the sum of pressure gradients over topography (the so-called mountain torque), surface friction, and torque due to topographic gravity waves \citep{Lejenas1997}. 
Appropriate control variables could be the surface pressure, orographic gravity wave drag parameters, or surface friction. 
We defer exploration of this idea to future research. 


