
%-------------------WACCM-------------------------
WACCM \citep{Marsh2013} extends the dynamical core of CAM up to $5 \times 10^{-6}$hPa (circa 140km, in the lower thermosphere) with 66 levels, and includes additional chemical and physical processes for these regions. 
The experiments here use WACCM version 4 \hl{with no interactive chemistry}.
WACCM's three highest model levels (starting from about $2 \times 10^{-5}$hPa) have strong horizontal diffusion to absorb planetary waves. 
%% Kevin Reader's answer about the WACCM sponge layers: 
%The number of damped levels depends on the dissipation choice in the WACCM atm_in namelist variable div24del2flag. For value 2 (ldiv2 diffusion) there are 3 damped levels. For values 4 and 42 (ldiv4 and ldiv4+ldel2) there are 4 damped levels.
%In addition, the advection operators go to first order for the top 1/8 of the levels, which is 8 for the 66 level model.  I believe that this is a gentler
The DART-WACCM simulations in this study were also run at $1.9^{\text{o}} \times 2.5^{\text{o}}$ horizontal resolution, with 66 vertical hybrid coordinate levels. 
%dissipation, so I think that we should set highest_state_pressure_Pa so that only levels 1-3 (or 4) feel no observations.

%-------------------experiments-------------------------
Interfaces have been developed between DART and both CAM \citep{Raeder2012} and WACCM \citep{Pedatella2013}.
The experiments used in this study are summarized in Table \ref{tab:expts}.

%-----WACCM experiments
\subsubsection{DART-WACCM Experiments}
Experiments \WACCMNODA-\WACCMGLOBAL~ \hl{were developed for a different study (in preparation)}, and assimilate a range of real meteorological observations over the winter of 2009-10. 
In Section \ref{sec:evaluation_variable}, we use these experiments  
to examine how Earth rotation parameters reflect the constraint imposed by other observations in a data assimilation system.  
Experiment \WACCMNODA~runs a 40-member ensemble of WACCM simulations forward for six months starting on 1 Oct 2009, with no assimilation. 
\WACCMTROPICS~and \WACCMGLOBAL~use the same initial ensemble, but assimilate the meteorological observations that are also used in the NCEP/NCAR reanalysis, i.e. winds and temperatures from radiosondes and aircraft \citep{Saha2010}, as well as refractivity profiles measured via GPS radio-occultation by the COSMIC satellite constellation \citep{Anthes2008}.
The observations extend from the surface to about 2 hPa. 
\WACCMTROPICS~assimilates the observations only in the 30S-30N tropical band, while \WACCMGLOBAL~assimilates over the entire globe.
Though experiments \WACCMNODA-\WACCMGLOBAL~were run for six months, only the first 60 days of assimilation are discussed in this study.  

Experiments \WACCMTROPICS-\WACCMGLOBAL~both use a Gaspari-Cohn function \citep{Gaspari1999} to localize observations, with a half-width of 0.2 radians in the horizontal, and 0.5 scale heights in the vertical.  
Adaptive ensemble inflation \citep{Anderson2009tellus} is used to increase the ensemble spread when necessary; the inflation factor varies in space and time and is proportional to the distance between the ensemble mean and the observation, given the uncertainties of each.  
No analysis increment is allowed above 0.1 hPa, in order to prevent the generation of spurious gravity waves resulting from a large ensemble spread in the absence of upper level observations. 
\hl{Cite paper by Saroja here?}

%%---localization and inflation in Nick's 2014 paper: 
%Similar to Pedatella et al. [2013], the horizontal and vertical localizations are specified with a Gaspari-Cohn function [Gaspari and Cohn, 1999]
%with half widths of 0.2 radians and 0.5 in ln(p∕p0) coordinates, respectively. Spatially and temporally varying adaptive inflation is used to inflate the ensemble prior to the assimilation in order to prevent collapse of the model spread [Anderson, 2009].

%%---localization and inflation in Nick's 2013 paper: 
%The most notable change is that we specify the ver- tical localization using a Gaspari-Cohn function [Gaspari and Cohn, 1999] with a half width of 0.5 in ln(p/p0) coor- dinates, where p is pressure and p0 is the surface pressure. This differs from the vertical localization of 1000 hPa in p used by Raeder et al. [2012]. 

