Using observed LOD data to judge the overall fidelity of an atmosphere model: 
Boer1990.

Using observed LOD data to evaluate reanalysis datasets: Yu1999, Ayoma2000, 
Paek2012

It is still difficult to reconcile  differences between the seasonal 
polar wobble and the angular momentum fluctuations predicted by different 
components of the Earth system: Dobslaw2010.  

Nastula2009 used statistical analysis and an atmosphere model to estimate which 
regions cause the observed polar wobbles on various timescales.  


Many previous studies have used Earth rotation data to evaluat the accuracy of 
and identify biases in atmosphere models \citep{Boer1990, Rosen2000}, ocean 
models \citep{Gross1996a} or reanalysis data \citep{Yu1999, Ayoma2000, 
Paek2012},
%Ayoma2000 identified potential problems in reanalysis wind data due to 
discrepancies between predicted and observed LOD variations.
%Paek2012 used LOD observations to evaluate the accuracy of and identify 
biases in various reanalysis sets.
%Rosen2000 used observed LOD anomalies to evaluate the interannual 
variability simulated by AMIP models.
% Gross2006a used observed polar motion to identify discrepancies in ocean 
tide models: 


------------------------------------------------------------

@article{Yu1999,
author = {Yu, N and Zheng, D and Wu, H},
file = {:home/lneef/Dropbox/NATHAN/Reprints/yuetal1999.pdf:pdf},
journal = {J. Geod.},
keywords = {length-of-day variation \'{a} atmospheric},
mendeley-groups = {Earth Rotation},
pages = {385--390},
title = {{Contribution of new AAM data source to DLOD excitation}},
volume = {73},
year = {1999}
}

@article{Gross1996a,
author = {Gross, Richard S and Hamdan, Kamal H and Boggs, Dale H},
file = {:home/lneef/Dropbox/NATHAN/Reprints/grossetal1996a.pdf:pdf},
journal = {Geophys. Res. Lett.},
mendeley-groups = {Earth Rotation/Polar Motion},
number = {14},
pages = {1809--1812},
title = {{Evidence for excitation of polar motion by fortnightly ocean tides}},
volume = {23},
year = {1996}
}

@article{Boer1990,
author = {Boer, G.J.},
file = {:home/lneef/Dropbox/NATHAN/Reprints/boer1990.pdf:pdf},
journal = {J. Geophys. Res.},
mendeley-groups = {Earth Rotation,Earth Rotation/Torques},
number = {D5},
pages = {5511--5531},
title = {{Earth-Atmosphere Exchange of Angular Momentum Simulated in a General 
Circulation Model and Implications for the Length of Day}},
volume = {95},
year = {1990}
}

@article{Aoyama2000,
author = {Aoyama, Yuichi and Naito, Isao},
doi = {10.1029/2000JD900101},
file = {:home/lneef/.local/share/data/Mendeley Ltd./Mendeley 
Desktop/Downloaded/Aoyama, Naito - 2000 - Wind contributions to the Earth's 
angular momentum budgets in seasonal variation.pdf:pdf},
issn = {0148-0227},
journal = {J. Geophys. Res.},
mendeley-groups = {Earth Rotation},
month = may,
number = {D10},
pages = {12417},
title = {{Wind contributions to the Earth's angular momentum budgets in seasonal 
variation}},
url = {http://www.agu.org/pubs/crossref/2000/2000JD900101.shtml},
volume = {105},
year = {2000}
}

@article{Dobslaw2010,
author = {Dobslaw, H. and Dill, R. and Gr\"{o}tzsch, a. and Brzeziński, a. and 
Thomas, M.},
doi = {10.1029/2009JB007127},
file = {:home/lneef/.local/share/data/Mendeley Ltd./Mendeley 
Desktop/Downloaded/Dobslaw et al. - 2010 - Seasonal polar motion excitation from 
numerical models of atmosphere, ocean, and continental hydrosphere.pdf:pdf},
issn = {0148-0227},
journal = {J. Geophys. Res.},
mendeley-groups = {Earth Rotation,Earth Rotation/Polar Motion},
month = oct,
number = {B10},
pages = {1--11},
title = {{Seasonal polar motion excitation from numerical models of atmosphere, 
ocean, and continental hydrosphere}},
url = {http://www.agu.org/pubs/crossref/2010/2009JB007127.shtml},
volume = {115},
year = {2010}
}

@article{Paek2012a,
author = {Paek, Houk and Huang, Huei-Ping},
doi = {10.1029/2012JD018105},
issn = {0148-0227},
journal = {J. Geophys. Res.},
keywords = {angular momentum,length of day,reanalysis},
mendeley-groups = {Earth Rotation,Reanalyses,Earth Rotation/LOD},
month = dec,
number = {D20},
pages = {121207115945000},
title = {{A comparison of the interannual variability in atmospheric angular 
momentum and length-of-day using multiple reanalysis data sets}},
url = {http://www.agu.org/pubs/crossref/2012/2012JD018105.shtml 
http://journals.ametsoc.org/doi/abs/10.1175/JCLI-D-12-00515.1},
volume = {117},
year = {2012}
}


@article{Nastula2009,
annote = {Which geographic regions are the main sources of excitation for polar
author = {Nastula, J. and Salstein, D. and Kolaczek, B.},
doi = {10.1029/2008JB005605},
file = {:home/lneef/.local/share/data/Mendeley Ltd./Mendeley 
Desktop/Downloaded/Nastula, Salstein, Kolaczek - 2009 - Patterns of atmospheric 
excitation functions of polar motion from high‐resolution regional 
sectors.pdf:pdf},
issn = {0148-0227},
journal = {J. Geophys. Res.},
mendeley-groups = {Earth Rotation},
month = apr,
number = {B4},
pages = {1--17},
title = {{Patterns of atmospheric excitation functions of polar motion from 
high‐resolution regional sectors}},
url = {http://www.agu.org/pubs/crossref/2009/2008JB005605.shtml},
volume = {114},
year = {2009}
}

@article{Rosen2000,
author = {Rosen, R. D. and Salstein, D. a.},
doi = {10.1007/s003820000076},
file = {:home/lneef/.local/share/data/Mendeley Ltd./Mendeley 
Desktop/Downloaded/Rosen, Salstein - 2000 - Multidecadal signals in the 
interannual variability of atmospheric angular momentum.pdf:pdf},
issn = {0930-7575},
journal = {Clim. Dyn.},
mendeley-groups = {Earth Rotation,Earth Rotation/LOD},
month = sep,
number = {9},
pages = {693--700},
title = {{Multidecadal signals in the interannual variability of atmospheric 
angular momentum}},
url = 
{
http://www.springerlink.com/openurl.asp?genre=article\&id=doi:10.1007/s003820000
076},
volume = {16},
year = {2000}
}

